Comments on Latent Ranking

### 2020-08-26 output_scripts/aug_updates.Rmd
start adding updates for James

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
library(ggplot2)
library(gridExtra)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

@@

To see if closer ranked mice fight more, we would like to look
at the conditional intensity as a function of the absolute difference in the
latent ranks.

There are two components to this. Whether a pair of mice will be in the
active state (where fighting is more likely), and, in this active state,
the value of the excitation parameter (which controls the likelihood of
fighting again following a previous fight).

```{r difference in ranking for all cohorts}
for(current_cohort in 1:10) {
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
  # have 1000 draws here
  eta_1_mean <- mean(sim_cohort_mmhp$eta_1)
  eta_2_mean <- mean(sim_cohort_mmhp$eta_2)
  eta_3_mean <- mean(sim_cohort_mmhp$eta_3)

  f_1 <- seq(from = 0.01, to = 0.99, by = 0.01)
  f_2 <- seq(from = 0.01, to = 0.99, by = 0.01)

  possible_vals <- expand_grid(f_1,f_2)

  excitation <- map2_dbl(possible_vals$f_1, possible_vals$f_2,function(.x,.y) 
      eta_1_mean*.x*.y*exp(-eta_2_mean*abs(.x-.y)) )

  prop_active <- map2_dbl(possible_vals$f_1, possible_vals$f_2,function(.x,.y) 
      1/(1+exp(-eta_3_mean*(.x-.y)))
     )

  possible_vals$excit <- excitation
  possible_vals$active <- prop_active

  plot_alpha <- possible_vals %>% ggplot(aes(f_1,f_2)) + 
    geom_raster(aes(fill=excit)) + 
    scale_fill_continuous(name = "alpha",low = "red", high = "blue") + 
    ggtitle(paste("Cohort ", current_cohort,
                  " excitation parameter",sep=""))

  plot_active <- possible_vals %>% ggplot(aes(f_1,f_2)) + 
    geom_raster(aes(fill=active)) + 
    scale_fill_continuous(name = "Prop. active",low = "red", high = "blue") + 
    ggtitle(paste("Cohort ", current_cohort,
                  " proportion of time active",sep=""))
  print(plot_alpha)
  print(plot_active)
}
```

```{r intensity as function of abs diff in ranks, include=FALSE}
current_cohort <- 1
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
  # have 1000 draws here
eta_1_mean <- mean(sim_cohort_mmhp$eta_1)
eta_2_mean <- mean(sim_cohort_mmhp$eta_2)
eta_3_mean <- mean(sim_cohort_mmhp$eta_3)

f_1 <- seq(from = 0.01, to = 0.99, by = 0.01)
f_2 <- seq(from = 0.01, to = 0.99, by = 0.01)

possible_vals <- expand_grid(f_1,f_2)

excitation <- map2_dbl(possible_vals$f_1, possible_vals$f_2,function(.x,.y) 
    eta_1_mean*.x*.y*exp(-eta_2_mean*abs(.x-.y)) )

prop_active <- map2_dbl(possible_vals$f_1, possible_vals$f_2,function(.x,.y) 
    1/(1+exp(-eta_3_mean*(.x-.y)))
   )

possible_vals$excit <- excitation
possible_vals$active <- prop_active

possible_vals %>% mutate(abs_diff = abs(f_1-f_2)) %>%
  ggplot(aes(abs_diff,excit)) + geom_point()

possible_vals %>% mutate(abs_diff = abs(f_1-f_2)) %>%
  ggplot(aes(abs_diff,prop_active)) + geom_point() +
  xlab("Absolute Difference in Ranks") +
  ylab("Proportion of Active Time")

# plot_alpha <- possible_vals %>% ggplot(aes(f_1,f_2)) + 
#   geom_raster(aes(fill=excit)) + 
#   scale_fill_continuous(name = "alpha",low = "red", high = "blue") + 
#   ggtitle(paste("Cohort ", current_cohort,sep=""))
```

Based on these plots, pairs of mice are more likely to be in the active state
when the instigating mouse is ranked significantly higher than the mouse it
fights. However, in the active state, the excitation parameter (and
the likelihood of continuing to fight) is higher 
when both mice are higher ranked. This is relatively consistent across each of
the cohorts here.

# Exploring $\eta$ parameters

The parameters $\eta_1,\eta_2,\eta_3$ in this model are related to the
probability of being in the active state and the excitation parameter 
when in this active state. We can first look at the posterior distribution
of these parameters and their posterior means.

```{r posterior pairs for eta, message=FALSE}
for(current_cohort in 1:10) {
  #cat("Cohort, ", current_cohort,"\n",sep="")
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  eta <- tibble(eta_1 = sim_cohort_mmhp$eta_1,
                eta_2 = sim_cohort_mmhp$eta_2,
                eta_3 = sim_cohort_mmhp$eta_3)
  print("Posterior Mean of eta parameters")
  print( eta %>% map_dbl(mean) )

  p <- GGally::ggpairs(eta) + 
    ggtitle(paste("Cohort ", current_cohort,sep=""))

@@

$$
\lambda^{i,j}(t) = \gamma_i + \zeta_j +(\lambda_1-\gamma_i-\zeta_j)Z^{i,j}(t) 
$$

$$
+
\eta_1 f_i f_j 
\exp(-\eta_2|f_i-f_j|)Z^{i,j}(t)\sum_{k}\exp(-\beta(t-t_{k}^{i,j})),
$$
with
$$
q_1^{i,j} = \exp(-\eta_3 f_i),\ q_0^{i,j} = \exp(-\eta_3 f_j).
$$
So here, these parameters correspond to:

- $\eta_1$ excitation as a function of the product of the latent ranks.
- $\eta_2$ is related to the absolute difference between the ranks of the mice. 
The posterior mean for this parameter is less than 1 for each cohort, and
the absolute difference between the ranks is upper bounded by 1. Larger
values of $\eta_1$ and small values of $\eta_2$ correspond to a larger
excitation parameter, with pairs more likely to continue fighting.
- $\eta_3$ is related to the state transition probabilities. As the rankings
are between zero and 1, larger values of $\eta_3$ correspond to a higher 
likelihood of being in the active state, particularly if one mice has a high or
low latent ranking.

@@

```{r same plot from previous}
for(current_cohort in 1:10) {
  #cat("Cohort ", current_cohort,"\n",sep="")
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  mean_latent_rank <- apply(sim_cohort_mmhp$f,2,mean)

  out_degree <- apply(sim_cohort_mmhp$gamma,2,mean)
  in_degree <- apply(sim_cohort_mmhp$zeta,2,mean)

  degs <- tibble(expert = expert_rank_10[[current_cohort]],
                  latent = mean_latent_rank,#[expert],
                 out_deg = out_degree,#[expert],
                 in_deg = in_degree)#[expert])

  p <- ggplot(degs,aes(latent,out_deg)) + geom_point() + 
    #scale_y_continuous(breaks = c(12:1), trans = "reverse") + 
    ggtitle(paste("Out Degree vs Latent Cohort ", current_cohort,sep="")) + 
    ylab("Out Degree (Gamma)") + xlab("Latent Rank")
  #print(p)
  q <- ggplot(degs,aes(latent,in_deg)) + geom_point() +
    ggtitle(paste("In Degree vs Latent Cohort ",current_cohort,sep="")) +
    ylab("In Degree (Zeta)") + xlab("Latent Rank")
  #print(q)
  grid.arrange(p,q, nrow = 1)
}

```

There is a weak relationship between the out degree 
(baseline rate of starting a fight) and the latent rank,
with an even weaker relationship between the in degree
(baseline rate of being fought) and the latent rank.

<!-- # Shuffling the data -->

>>>>

library(rstan)
Sys.setenv(LOCAL_CPPFLAGS = '-march=corei7 -mtune=corei7')
options(mc.cores = parallel::detectCores())
library(tidyverse)
library(R.utils)
library(fields)
library(bayesplot)
library(compete)
library(RColorBrewer)
library(ggplot2)
source(paste(func_dir,'naiveRankHierarchy.R',sep = ""))
source(paste(func_dir,'expertRankHierarchy.R',sep = ""))
source(paste(func_dir,'cleanData.R',sep = "")) 
source(paste(func_dir,'prepareDataStan.R',sep = "")) 
source(paste(func_dir,'inferLatentMmhp.R',sep = ""))
source(paste(func_dir,'plotUtil.R',sep = "")) 
source(paste(func_dir,'mmhp.R',sep = ""))
source(paste(func_dir,'uniHawkes.R',sep = ""))
source(paste(func_dir,'simulatePrediction.R',sep = ""))
source(paste(func_dir,'myGlicko.R',sep = ""))
source(paste(func_dir,'matrixPlotParameter.R',sep = ""))
source(paste(func_dir,'residualStructureScore.R',sep = ""))
source("https://gist.githubusercontent.com/jalapic/6ca3ece44bdcdc522bb735f183aa0ca0/raw/1a07f469eff08117121b6cbebcd22fb7569e3ee8/compete_extra.R")

@@

# Exploring eta parameters

```{r}
for(current_cohort in 1:10) {
  #cat("Cohort, ", current_cohort,"\n",sep="")
  load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))

  eta <- tibble(eta_1 = sim_cohort_mmhp$eta_1,
                eta_2 = sim_cohort_mmhp$eta_2,
                eta_3 = sim_cohort_mmhp$eta_3)
  p <- GGally::ggpairs(eta) + 
    ggtitle(paste("Cohort ", current_cohort,sep=""))

@@

$$
\lambda^{i,j}(t) = \gamma_i + \zeta_j +(\lambda_1-\gamma_i-\zeta_j)Z^{i,j}(t) +
\eta_1 f_i f_j 
\exp(-\eta_2|f_i-f_j|)Z^{i,j}(t)\sum_{k}\exp(-\beta(t-t_{k}^{i,j})),
$$
where 
$$
q_1^{i,j} = \exp(-\eta_3 f_i),\ q_0^{i,j} = \exp(-\eta_3 f_j).
$$
So here, these parameters correspond to:
- $\eta_1$

@@
